Boosting Higgs Pair Production in the $bb\bar{b}\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-07-08 paper/sec-conclusions.tex
work on paper

>>>>

Finally, it would be important to estimate the prospects for
Higgs pair production in the $4b$ at a Future Circular
Collider with a center of mass energy of 100 TeV, along the lines
of similar studies~\cite{Barr:2014sga}.
%
While signal yields would be certainly increased, also the QCD
multijet background would grow, so it is necessary to repeat
our analysis, including full background simulation, for 100 TeV,
in order to be able to draw any definite conclusion here.

### 2015-07-08 paper/sec-mva.tex
work on paper

  This variable is defined by reclustering the constituents of a jet with the
  $k_t$ algorithm, which usually clusters last the harder constituents, and then
  taking the $k_t$ distance measure between the two subjets at the final stage of the recombination
  \be
\sqrt{d_{12}} \equiv {\rm min}\lp p_{T,1},p_{T,2}\rp \cdot \Delta R_{12} \, ,
\ee
where one can also define related variables like $\sqrt{d_{23}}$ that corresponds to the
splitting scale in the second-to-last clustering.

@@

  The $N$-subjettiness variables $\tau_N$ are defined by clustering the constituents
  of a jet with the exclusive $k_t$ algorithm and requiring that $N$ subjets are found,
  \be
  \tau_N \equiv \frac{1}{d_0} \sum_k p_{T,k}\cdot {\rm min}\lp \delta R_{1k}, \ldots,
  \delta R_{Nk}\rp \, , \qquad d_0\equiv \sum_k p_{T,k}\cdot R \, ,
  \ee
  with $ p_{T,k}$ the $p_T$ of the constituent $k$ and $\delta R_{ik}$ the distance from
  subjet $i$ to constituent $k$.
  %
  In this work in particular we use as input to the MVA
  \be
\tau_{21} \equiv \frac{\tau_2}{\tau_1} \, ,
  \ee
  which provide good discrimination between QCD jets and jets arising from the decay of
  a heavy resonance.

@@

  The energy correlation double ratio $C_2$ is defined
  as
  \be
C_2^{\beta} \equiv \frac{ {\rm ECF}(3,\beta) {\rm ECF}(1,\beta)}{{\rm ECF}(2,\beta)} \, ,
  \ee
  where the energy correlation functions ${\rm ECF}(3,\beta)$ are defined
  in~\cite{Larkoski:2013eya} with the motivation that $(N+1)$-point correlators
  are sensitive to $N$-prong substructure.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Boosted category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm mva}=0$} & \multicolumn{4}{c}{$y_{\rm mva}=0.85$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      193   &  $3.8\,10^4$    &  0.99    &  $5.1\,10^{-3}$  &  &  &  & \\
        \hline
  \end{tabular}
   $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Intermediate category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm mva}=0$} & \multicolumn{4}{c}{$y_{\rm mva}=0.85$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      660   &   $2.2\,10^6$   &   0.45   &  $3.1\,10^{-4}$ &  &  &  & \\
        \hline
  \end{tabular}
    $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Resolved category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm mva}=0$} & \multicolumn{4}{c}{$y_{\rm mva}=0.85$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      $1.6\,10^{3}$   &   $1.1\,10^{7}$   & 0.50     &  $1.5\,10^{-4}$   &  &  &  & \\
        \hline
  \end{tabular}
  \caption{\small The results for the number of signal and
    background events
    at the HL-LHC for both the cut-based analysis and after the
    optimal MVA cut, as well as the $S/\sqrt{B}$ and $S/B$
    values in the two cases.
    %
    From top to bottom we show the results for boosted,
    intermediate and resolved categories.
    %
    Note that the optimal value of the MVA cut
    is different in each of the three categories.
    \label{table:cutflowMVA}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Another look at jet substructure variables}

It is now worth trying to understand where does the increase in signal significance
achieved after the MVA comes from.
%
As we will show now, the MVA has correctly identified which of the substructure
variables are more discriminant between signal and background distributions,
and combined this information to determine an optimal cut to maximize signal
significance.

For instance, in the boosted category, the analyses of the
size of the neural network weights clearly indicates that the subjetiness
$\tau_{12}$ and the energy correlations $C_2$ are two of the most useful
variables.
%
To verify that indeed these two variables are quite different between signal
and background events, in Fig.~\ref{fig:mva_substructure_1}
we show the subjetiness $\tau_{12}$ and the energy correlation between
signal and background.
%
We indeed verify that the two shapes are quite different,
reflecting the differences in the internal structure of jets
between QCD jets and jets originated from the Higgs decays.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.48\textwidth]{plots/tau21_h0_res_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/EEC_C2_h0_res_C1_boost.pdf}
  \caption{\small Left plot: subjetiness.
    Right plot: energy correlations.
}
\label{fig:mva_substructure_1}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

  This variable is defined by reclustering the constituents of a jet with the
  $k_t$ algorithm, which usually clusters last the harder constituents, and then
  taking the $k_t$ distance measure between the two subjets at the final stage of the recombination
  \be
  \label{eq:ktsplitting}
\sqrt{d_{12}} \equiv {\rm min}\lp p_{T,1},p_{T,2}\rp \cdot \Delta R_{12} \, ,
\ee
where one can also define related variables like $\sqrt{d_{23}}$ that corresponds to the
splitting scale in the second-to-last clustering.

@@

  The $N$-subjettiness variables $\tau_N$ are defined by clustering the constituents
  of a jet with the exclusive $k_t$ algorithm and requiring that $N$ subjets are found,
  \be
  \tau_N \equiv \frac{1}{d_0} \sum_k p_{T,k}\cdot {\rm min}\lp \delta R_{1k}, \ldots,
  \delta R_{Nk}\rp \, , \qquad d_0\equiv \sum_k p_{T,k}\cdot R \, ,
  \ee
  with $ p_{T,k}$ the $p_T$ of the constituent $k$ and $\delta R_{ik}$ the distance from
  subjet $i$ to constituent $k$.
  %
  In this work in particular we use as input to the MVA
  \be
  \label{eq:tau21}
\tau_{21} \equiv \frac{\tau_2}{\tau_1} \, ,
  \ee
  which provide good discrimination between QCD jets and jets arising from the decay of
  a heavy resonance.

@@

  The energy correlation double ratio $C_2$ is defined
  as
  \be
  \label{eq:c2}
C_2^{\beta} \equiv \frac{ {\rm ECF}(3,\beta) {\rm ECF}(1,\beta)}{{\rm ECF}(2,\beta)} \, ,
  \ee
  where the energy correlation functions ${\rm ECF}(3,\beta)$ are defined
  in~\cite{Larkoski:2013eya} with the motivation that $(N+1)$-point correlators
  are sensitive to $N$-prong substructure.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Boosted category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm mva}=0$} & \multicolumn{4}{c}{$y_{\rm mva}=0.86$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      193   &  $3.8\,10^4$    &  0.99    &  $5.1\,10^{-3}$  & 107 & 1040 & 3.31  & 0.10\\
        \hline
  \end{tabular}
   $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Intermediate category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm mva}=0$} & \multicolumn{4}{c}{$y_{\rm mva}=0.49$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
      660   &   $2.2\,10^6$   &   0.45   &  $3.1\,10^{-4}$ & 480  & $3.2\,10^5$ & 0.85  & $1.5\,10^{-3}$ \\
        \hline
  \end{tabular}
    $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Resolved category}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm mva}=0$} & \multicolumn{4}{c}{$y_{\rm mva}=0.17$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
    $1.6\,10^{3}$   &   $1.1\,10^{7}$   & 0.50     &  $1.5\,10^{-4}$   & 968 &
    $1.2\,10^{6}$  & 0.90 & $8.2\,10^{-4}$ \\
        \hline
  \end{tabular}
  \caption{\small The results for the number of signal and
    background events
    at the HL-LHC for both the cut-based analysis and after the
    optimal MVA cut, as well as the $S/\sqrt{B}$ and $S/B$
    values in the two cases.
    %
    From top to bottom we show the results for boosted,
    intermediate and resolved categories.
    %
    Note that the optimal value of the MVA cut
    is different in each of the three categories.
    \label{table:cutflowMVA}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

From Table~\ref{table:cutflowMVA} we see that after the
MVA the signal significance in the boosted category increases
from 0.99 to 3.31, with $S/B$ increasing from $0.005$ to $0.10$,
with still more that 100 signal events at the HL-LHC.
%
Table~\ref{table:cutflowMVA} also shows that for the intermediate
and resolved categories, the improvement due to the use
of the MVA is rather moderate, and signal significance is always
small since backgrounds are still very large.

Table~\ref{fig:sb_mva} contains the main result of this paper:
at the HL-LHC, it should be possible to claim evidence of
Higgs pair production using only the $4b$ final state.
%
Needless to say, to claim discovery and to improve the accuracy
of the measurement of the trilinear coupling it would still
be needed to combine this channel with others like
$b\bar{b}\gamma\gamma$ and $b\bar{b}\tau\tau$, but in any case
it is always important to have solid evidence from
individual channels.

From Table~\ref{fig:sb_mva} it is also possible to
extract the signal significance that would be obtained
at the end of Run II with $\mathcal{L}_{\rm int}=300$ fb$^{-1}$:
in the boosted category we expect 11 signal events and 100 background
events, leading to a $S\sqrt{B}=1.1$, too small specially accounting
for realistic systematic experimental uncertainties.
%
So the full luminosity of the HL-LHC is really needed to exploit
the information contained in this final state.

In the next section we discuss how the expectations for the HL-LHC
summarized in  Table~\ref{fig:sb_mva} are modified in different
conservative and aggressive scenarios for the performance of the
ATLAS and CMS experiments, in particular how the results depend
on the $b$-tagging performance and in the detector momentum
resolution.
%
But before that let's illustrate how the improvement in $S/\sqrt{B}$
obtained from the MVA arises from the power of the jet
substructure variables.

\subsection{Another look at the jet substructure variables}

The power of the MVA applied to the boosted category lies
on the information contained on the various substructure variables
used.
%
From the distribution of total associated ANN weights in the
boosted category shown in Fig.~\ref{fig:nnweights} we see that
all substructure variables contain useful information,
but specially the $k_t$ splitting scale $\sqrt{d_{12}}$ of the
subleading Higgs candidate, the $\tau_{21}$ subjettiness variable for
the leading Higgs
and the ratio of energy correlators $D_2$ of both the leading
and subleading Higgs candidates.
%
Interestingly, other substructure variables like $\tau_{21}$ for
the subleading Higgs seem to contain less information (or perhaps
redundant as compared to the other variables).

The information contained Fig.~\ref{fig:nnweights} should be
reflected in the corresponding distributions of these substructure
variables: those with a higher discrimination power should show
clear differences in shape between the signal and background events,
while those that carry less information should have similar
shapes for signal and background.
%
There is also the possibility of redundant variables, where signal
and background are quite different but the same information
is already provided by other input variables.

In Fig.~\ref{fig:mva_substructure_1}
we show these substructure variables: the
$k_t$ splitting scale $\sqrt{d_{12}}$, Eq.~(\ref{eq:ktsplitting}) for
the subleading Higgs, the subjettiness ratio $\tau_{12}$
Eq.~(\ref{eq:tau21})
for the leading Higgs,
and the energy correlation double ratio $D_2$
Eq.~\ref{eq:c2}
for the leading and subleading Higgs.
%
We can indeed verify that the shapes of the signal
and background distributions are quite different,
reflecting the differences in the internal structure of jets
between QCD jets and jets originated from the Higgs decays.
%
This is true for all variables considered here.
%
As before, all the distributions in
Fig.~\ref{fig:mva_substructure_1}
have been normalized to unity.
%
All the distributions shown in this section refer to the boosted category.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.48\textwidth]{plots/split12_H1_res_C1_boost.pdf} 
  \includegraphics[width=0.48\textwidth]{plots/tau21_h0_res_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/EEC_C2_h0_res_C1_boost.pdf}
  \includegraphics[width=0.48\textwidth]{plots/EEC_C2_h0_res_C1_boost.pdf}
  \caption{\small Distribution of substructure variables
    in the boosted category at the end of the cut-based
    analysis, used as input to the MVA.
    %
    From top to bottom and from left to right we show  the
$k_t$ splitting scale for
the subleading Higgs, the subjettiness ratio $\tau_{12}$
for the leading Higgs,
and the energy correlation double ratio $D_2$
for the leading and subleading Higgs.
%
All distributions have been normalized to unity.
}
\label{fig:mva_substructure_1}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Finally, let us show the distributions of those kinematic variables
that according to the MVA provide a small amount of discrimination
between signal and background, either because there is no real information or
because it is redundant.

{\bf add missing histograms}
